---
title: "Training course 2024"
author: "Emmanuel L Mayenga"
date: "2024-11-18"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# course part 1 - Laboratory

# we selected 20 isolates

- 6 are *Klebsiella spp* isolates
- 6 are *E. coli* isolates
- 8 are **Salmonella spp** isolates

 <!-- 
 TODO: I am a comments
 damned lab work it failed too much today
-->


## Installation  of packages


```{r}
#install.packages("here")
#install.packages("Diagrammer")
```
 ctrl + alt + I (make a cell - shortcut on the keyboard)

```{r}
angry_or_not <- "not"
```



 give content of the object (container)
```{r}
 angry_or_not
```


Emmanuel, are you angry or not? `r print(angry_or_not)`

# working with my data


```{r}
 here::here()
```


```{r}
path_data <-here::here("data", "2024-09-25_INIKA_SAMPLING_ANALYSIS_HUMAN.csv")

path_data
```


```{r}
 human_data <-
  readr::read_csv2(path_data)
```


# Day 3

## Packages into memory 

```{r}
library(here)
library(tidyverse)
```
F1


```{r}
?library()
help("library")
```


## Exploration of a data frame

- Names of columns

```{r}
colnames(human_data)
```


- Overview of data and data types
```{r}
dput(colnames(human_data))
```


- explanation: c() is to make vectors
```{r}
test <- c("one", "2", 3)

test 
typeof(test)

class(test)
```

```{r}
test2 <-c(1,2,3)
typeof(test2)
class(test2)
```

```{r}
test <- c("one", "2", 3)
test
test[3]
typeof(test[3])
typeof(test)
```

``{r}
test <- c(1, 2, 3)
test
typeof(test2)
class(test)

```{r}
test2*test2
```

- renaming columns names (easier to use)


```{r}
original_names <- colnames(human_data)
original_names

```


```{r}
library(tidyverse)
new_names <- str_replace_all(original_names, " ", "_")
new_names
new_names[88]
```

- regex pattern
```{r}
new_names2 <- str_replace_all(
  new_names, "[?/,;.*()-]", "_") %>%
  str_remove_all("(^_*)|(_*$)")
new_names2
new_names2[88]
```

```{r}
colnames(human_data) <- new_names2
```


- selecting columns


```{r}
human_data_selection <- 
  human_data %>%
  select(INIKA_OH_TZ_ID, Age__yrs, Gender, Enter_a_date, Region, District, 
         Specify_if_other_district, Sample, Season, Origin_of_sample, 
         Which_class_grade_are_you, 
         Who_is_your_caretaker, 
         If_others__mention, 
         What_is_your_occupation_and_or_of_your_caretaker, 
         Have_you_ever_heard_about_AMR, If_yes__how_did_you_get_this_information, 
         Have_you_or_your_children_used_any_antibiotics_at_any_time, 
         If_yes__where_did_you_get_these_drugs_from,
         If_it_was_drug_sellers_or_pharmacy__did_you_have_a_prescription_from_the_doctor_prescriber,
         GPS_coordinates_latitude, GPS_coordinates_longitude) 

human_data_selection %>% View()
```
```{r}
glimpse(human_data_selection)
str(human_data_selection)
```

# HERE
- something does not make sense - empty column

```{r}
human_data_selection <-
  human_data_selection %>%
  select(-Specify_if_other_district) %>%
  mutate_at(vars(Age__yrs), as.integer) %>%
  mutate_if(is.character, factor) %>%
  mutate_at(vars(INIKA_OH_TZ_ID) , as.character)
```

Possible to use levels to get the different content of variables


```{r}
levels(human_data_selection$Region)
```


Summary
```{r}
summary(
  human_data_selection
)
```


- looking at the content of single columns
- different values contained coulmn
- 

- changing the types

```{r}
human_data_selection %>%
  filter(INIKA_OH_TZ_ID == "238")
```

```{r}
human_data_selection %>%
  filter(INIKA_OH_TZ_ID != "238") %>%
  filter(is.na(Gender))
# summary()

#%>%
# view()
```










